---
title: "Project_0525_HeartDisease"
author: "Anupma Mangla"
date: "5/25/2019"
output: html_document
---
## install the packages

```{r}
install.packages("gridExtra")
install.packages("data.table")
install.packages("caTools")
```

## Load the libraries

```{r}
library(tidyverse)
library(dplyr)
library(caTools)
library(gridExtra)
library(data.table)
```

# Load the dataset and look into the insights of data

```{r}
#heartData <- read.csv("~/Desktop/Project/heart.csv")
heartData <- read.csv("heart.csv")
input <- copy(heartData)
dim(input)
str(input)
```

# display the first 6 rows

```{r}
head(input)
```

# change the columns sex and target to their corresponding values
```{r}
input$sex <- ifelse((input$sex == 1), 'male','female')
input$target <- ifelse((input$target == 1), 'HeartDisease','No Heart Disease')

head(input)
```

# Number of observations: Healthy Vs Heart Disease cases

```{r}
table(input$target)
ggplot(input,aes(target, fill=target)) + geom_bar(stat="count") +
  ggtitle("Healthy Vs Heart Disease", subtitle="From Heart Disease UCI dataset ") + xlab("Target") + ylab("Number of People")
```

# Analyse heart disease risk factors: gender, age, high cholesterol, high blood pressure, diabetes 

# How does gender play a role in heart-disease?
### More males are healthy compared to females

```{r}

table(input$target, input$sex)
ggplot(input,aes(target, fill=target)) + geom_bar(stat="count") + ggtitle("Role of gender in Heart Disease") + xlab("Target") + ylab("Number of People") + facet_wrap(~sex, ncol=2,scale="fixed") 

```

# How does Age play a role in heart disease?
### We can see a lot more people in middle age with heart disease than the older age group, so we cant make judgement on age for heart disease

```{r}
# analazing heart disease over age distribution
p1 <- ggplot(data = input) + geom_bar(mapping = aes(x = age ,fill = target), position = "dodge") + labs(title = "Role of Age in heart disease") + xlab("Age distribution") + ylab("Number of people") + theme(legend.position = "top", legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))

p1
```

# How does cholestrol play a role in heart disease?
### we dont see many people with heart disease with very high cholestrol value, so high cholestrol does not seem to infer in heart disease

```{r}
## analazing heart disease over cholestoral
p2 <- ggplot(data = input) + geom_histogram(mapping = aes(x =  chol, fill = target), position = "dodge", bins = 15) + labs(title = "Role of Cholestrol in heart disease") + xlab("Cholestrol") + ylab("Number of people") + theme(legend.position = c(0.98, 0.98), legend.justification = c("right", "top"), legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))

p2
```

# How does resting BP play a role in heart disease?
###  High BP doesn't indicate higher number in heart disease patients

```{r}
# Role of resting BP 
p3 <- ggplot(data = input) + geom_histogram(mapping = aes(x =  trestbps, fill = target), position = "dodge", bins = 15) + labs(title = "Role of resting BP in heart disease") + xlab("Resting Blood Pressure") + ylab("Number of people") + theme( legend.position = c(0.98, 0.98), legend.justification = c("right", "top"), legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))

p3
```

# How does fasting blood sugar play a role in heart disease?
## we see more people with heart disease when sugar level is low, so high sugar does not seem to infer in heart disease
```{r}
# Role of fasting blood sugar
# fasting blood sugar > 120 mg/dl = 1, < 120 mg/dl = 0

p4 <- ggplot(data = input) + geom_histogram(mapping = aes(x =  fbs, fill = target), position = "dodge", bins =3) + labs(title = "Role of fasting blood sugar in heart disease", subtitle="Fasting blood sugar > 120 mg/dl (1= true)") + xlab("Fasting Blood Sugar") + ylab("Number of people") + theme( legend.position = c(0.98, 0.98), legend.justification = c("right", "top"), legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))
p4

```

# Create a multi-plot with all the 4 plots
```{r}

grid.arrange(p1,p2,p3,p4, nrow = 2, ncol = 2)

```

# How does max heart rate differ in healthy vs patients with heart disease?
##  Heart Disease patients have higher maximum heart rate than healthy patients
##  thalach, is the attribute to measure maximum heart rate achieved
  
```{r}
ggplot(input,aes(thalach, fill=target)) + 
  geom_histogram(aes(y=..density..),breaks=seq(70, 205, by=10), color="grey17") + 
  geom_density(alpha=.1, fill="black") +
  facet_wrap(~target, ncol=1,scale="fixed") +
  ggtitle("Max Heart Rate (thalach) Histogram") + xlab("Maximum Heart Rate Achieved") + ylab("Density / Count") + theme(legend.position ="top", legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))
```

# How does blood disorder Thalassemia play a role in heart disease?
## we can see thal type 2,3 have higher no. of people with heart disease

```{r}
## analazing heart disease over blood disorder Thalassemia (thal)
ggplot(data = input) + geom_bar(mapping = aes(x = thal ,fill = target), position = "dodge") + labs(title = "Role of blood disorder Thalassemia (thal) in presense of heart disease") + theme( legend.position = c(0.02, 0.98), legend.justification = c("left", "top"), legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))

```

# How does chest pain play a role in heart disease?
##  chest pain type 0,1,2 result in a higher number of patients with heart disease, we can infer that cp does play an important in heart disease

```{r}

ggplot(data = input) + geom_bar(mapping = aes(x = cp ,fill = target), position = "dodge") + labs(title = "Role of chest pain in heart disease") + xlab("Type of chest pain") + ylab("Number of people") + theme( legend.position = c(0.98, 0.98), legend.justification = c("right", "top"), legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))
```

# How does no. of blood vessels play a role in heart disease?
##  presence of less no. of blood vessels does indicate a higher chance in having heart disaese
```{r}
# analyzing heart disease over number of major blood vessels
ggplot(data = input) + geom_bar(mapping = aes(x =  ca, fill = target) , position = "dodge") +  labs(title = "Role of Major Blood Vessels (ca) in Heart disease") + xlab("Number of blood vessels in a person") + ylab("Number of people")  + theme( legend.position = c(0.98, 0.98), legend.justification = c("right", "top"), legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))

```

# What is the role of slope in predicting heart disease
##  With higher slope value, we can see higher number of people with heart disease

```{r}
# slope
ggplot(data = input) + geom_bar(mapping = aes(x = slope ,fill = target), position = "dodge") + labs(title = "The ST segment/heart rate slope as a predictor of CAD") + xlab("Slope") + ylab("Number of people") + theme( legend.position = c(0.99, 0.99), legend.justification = c("right", "top"), legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))
```

# Create Multiplot2

```{r}

# Chest pain type
p1 <- ggplot(data = input) + geom_bar(mapping = aes(x = cp ,fill = target), position = "dodge") + labs(title = "Role of chest pain in heart disease") + xlab("Type of chest pain") + ylab("Number of people") + theme( legend.position = "top", legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))

# thallasemia
p2 <- ggplot(data = input) + geom_bar(mapping = aes(x = thal ,fill = target), position = "dodge") + labs(title = "Role of blood disorder thalassemia") + xlab("Thalassemia") + ylab("Number of people") + theme( legend.position ="top", legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))

# Major blood vessels
p3 <- ggplot(data = input) + geom_bar(mapping = aes(x =  ca, fill = target) , position = "dodge") +  labs(title = "Role of Major Blood Vessels") + xlab("Number of blood vessels in a person") + ylab("Number of people")  + theme( legend.position = c(0.98, 0.98), legend.justification = c("right", "top"), legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))

# slope
p4 <- ggplot(data = input) + geom_bar(mapping = aes(x = slope ,fill = target), position = "dodge") + labs(title = "The ST segment/heart rate slope as a predictor of CAD") + xlab("Slope") + ylab("Number of people") + theme( legend.position = "top", legend.margin=margin(t = 0, unit='cm'), legend.key.size = unit(0.25, "cm"))

grid.arrange(p1,p2,p3,p4, nrow = 2, ncol = 2)

```

## how does gender and cholestrol combined together show the presence/absence of heart disease

```{r}
g<-ggplot(input)
g + geom_bar(stat="identity",aes(x=age,y=chol,fill=as.factor(sex)))+xlab("Age")+ylab("cholestrol")+labs(fill="Color Coding")
```

#After the Descriptive analysis, we create the model for predictive analysis

## we use the dataset to create the model using logistic regression to predict whether a person might or might not have heart disease based on the given attributes.

```{r}
heartData
```

## Split the data into training set 80% and testing test 20%
```{r}

split <- sample.split(heartData, SplitRatio = 0.8)
trainingSet <- subset(heartData, split == "TRUE")
testingSet <- subset(heartData, split == "FALSE")

nrow(trainingSet) 
nrow(testingSet)
nrow(heartData)

```

## Before creating the model, we munge the data and convert the categorical values to factors
```{r}
heartData$sex <- as.factor(heartData$sex)
heartData$cp <- factor(heartData$cp)
heartData$restecg <- factor(heartData$restecg)
heartData$fbs <- factor(heartData$fbs)
heartData$exang <- factor(heartData$exang)
heartData$ca <- factor(heartData$ca)
heartData$thal <- factor(heartData$thal)
heartData$target <- factor(heartData$target)

str(heartData)  ## structure of heartData

```

## Logistic regression is used to do the prediction analysis, glm(general linear model) function is used to model the data and training set is used as input
```{r}
## All the attributes are taken into consideration for analysis

logisticModel <- glm(target ~ age + sex + cp + trestbps + chol + fbs + restecg + thalach + exang + oldpeak + slope + ca + thal, data = trainingSet, family = 'binomial')

summary(logisticModel)

## use step function to eliminate the insignificant variables
newModel <- step(logisticModel, direction = "backward")
summary(newModel)
summary(table(heartData$age, heartData$sex, heartData$target, heartData$trestbps, heartData$cp, heartData$restecg))

```

## plot the model, visual display of regression residuals
```{r}
 plot(newModel)
```

## Run the testing data through the model, as logistic regression is a sigmoid function, if the value is greater than 0.5, we consider it as 1 else 0.
```{r}
result <- predict(newModel, newdata = testingSet, type='response')

result <- ifelse(result > 0.5, 1, 0)
```

# validate the model, compare the actual value of the result to the predicted value and determine the accuracy of the model
```{r}
linearPredCorrect <- data.frame(actual=testingSet$target, predicted=result, match = (testingSet$target == result))

linearPredCorrect
table(linearPredCorrect$match)

accuracy = length(linearPredCorrect$match[linearPredCorrect$match==TRUE])/count(testingSet) * 100
print(paste("Logistic regression has ", round(accuracy, 2), "% accuracy"))

```